---
title: "Reproducible Research: Peer Assessment 1"
author: "Austin Lo"
date: "April 4, 2016"
output: html_document
---
```{r, include=FALSE}
require(lattice)
require(ggplot2)
require(dplyr)
require(stringi)
require(stats)
require(gridExtra)
```

##Loading and preprocessing the data
```{r}

url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("repdata-data-activity.zip")) {
  download.file(url,"repdata-data-activity.zip")
  unzip("repdata-data-activity.zip")
}
ACT <- read.csv("activity.csv")
```
Calculate the Daily Sum
##What is mean total number of steps taken per day?
```{r }
ACT_Date_Sum<-ACT %>% group_by(date) %>% summarize(Total=sum(steps))
```
Calculate the Mean and Median of All Daily Sum 
```{r}
ACT_Date_Line=cbind(as.data.frame(rbind(mean(ACT_Date_Sum$Total,na.rm = T),median(ACT_Date_Sum$Total,na.rm = T))),Summary=c("Mean","Median"))
```
####<span style="color:blue">Answer : Mean Total Number of Steps Taken Per Day =  `r as.integer(mean(ACT_Date_Sum$Total,na.rm = T)) `</span>

```{r, echo=FALSE}
plot1<-ggplot(ACT_Date_Sum,aes(Total))+
       geom_histogram(bins=22,fill='#deebf7', color='#bdbdbd')+
       scale_y_continuous(breaks=seq(0, 10, 1))+
       labs(title="Histogram of Total Steps Taken Each Day",x="Total Steps")+
       geom_vline(data=ACT_Date_Line,mapping=aes(xintercept=V1,color=Summary))+
       geom_text(x=14000,y=9.5,label=paste("Mean=",round(ACT_Date_Line$V1[1],digits=0)))+
       geom_text(x=14000,y=9.0,label=paste("Median=",round(ACT_Date_Line$V1[2],digits=0)))
print(plot1)
```

##What is the average daily activity pattern?
Calculate the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) 
```{r}
ACT_Interval_Sum<-ACT %>% group_by(interval=as.factor(interval)) %>% summarize(Total=sum(steps,na.rm=T)/length(date))
ACT_Interval_Max<-ACT_Interval_Sum[which.max(ACT_Interval_Sum$Total),]
```

```{r echo=FALSE}
plot2<-ggplot(ACT_Interval_Sum,aes(x=interval,y=Total,group=1))+
       geom_line(color="steelblue")+
       labs(title="Average Steps of All Intervals",x="Intervals",y="Total Steps")+
       geom_text(x=10,y=ACT_Interval_Max$Total,hjust=0,
                 label=paste("Maximum Average Steps is",round(ACT_Interval_Max$Total,digits=2),"\nat ",
                 stri_sub(ACT_Interval_Max$interval,1,stri_length(ACT_Interval_Max$interval)-2),":",
                 stri_sub(ACT_Interval_Max$interval,-2,-1)))+
       theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(plot2)
```

##Imputing missing values
Imputed by Replacing the NA with Group Mean
```{r}
ACT_Imputed<-transform(ACT, steps = ifelse(is.na(steps), 
                       ave(steps, date, FUN = function(x) sum(x, na.rm = TRUE)/288), steps))
```
Calculate the Daily Sum of Imputed Data  
```{r}
ACT_Imputed_Date_Sum<-ACT_Imputed %>% group_by(date) %>% summarize(Total=sum(steps))
ACT_Imputed_Date_Line=cbind(as.data.frame(rbind(mean(ACT_Imputed_Date_Sum$Total,na.rm = T),
                                                median(ACT_Imputed_Date_Sum$Total,na.rm = T))),
                                                Summary=c("Mean","Median"))
```
####<span style="color:blue">Answer : Mean Total Number of Steps Taken Per Day(Imputed Data) =  `r as.integer(mean(ACT_Imputed_Date_Sum$Total,na.rm = T)) `</span>

```{r echo=FALSE}
plot3<-ggplot(ACT_Imputed_Date_Sum,aes(Total))+
  geom_histogram(bins=22,fill='#deebf7', color='#bdbdbd')+
  scale_y_continuous(breaks=seq(0, 10, 1))+
  labs(title="Histogram of Total Steps Taken Each Day(Imputed Dataset)",x="Total Steps")+
  geom_vline(data=ACT_Imputed_Date_Line,mapping=aes(xintercept=V1,color=Summary))+
  geom_text(x=14000,y=9.5,label=paste("Mean=",round(ACT_Imputed_Date_Line$V1[1],digits=0)))+
  geom_text(x=14000,y=9.0,label=paste("Median=",round(ACT_Imputed_Date_Line$V1[2],digits=0)))
print(plot3)
```

##Are there differences in activity patterns between weekdays and weekends?
Add Weekdays Column by Mutating from the date Column
```{r}
ACT_Weekday<-ACT_Imputed %>% mutate(weekdays = ifelse(weekdays(as.Date(date))=="Saturday"|weekdays(as.Date(date))=="Sunday","weekend","weekday"))
```
Group by Weekdays/Weekend and Calculate the Mean
```{r}
ACT_Weekday_group<-ACT_Weekday %>% group_by(interval,weekdays=as.factor(ACT_Weekday$weekdays))
ACT_Weekday_Sum<-as.data.frame(summarize(ACT_Weekday_group,Total=sum(steps)/length(date)))
```
####<span style="color:blue">Answer : The Average Steps between 5~9am and 18~19pm is higher in Weekdays, but the Average Steps between 9am~18pm is higher in Weekend</span>
```{r echo=FALSE}
plot4<-ggplot(ACT_Weekday_Sum,aes(interval, Total))+
              geom_line(aes(col=weekdays))+
              facet_grid(weekdays~.)+
              labs(title="Average Steps of All Intervals",x="Intervals",y="Total Steps")
print(plot4)  
```
